---
title: "Data clean up"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



## Deletions
Quarter
Marketing
Origin City Name
Origin State Name
Origin Wac
Dest Airport ID
Dest Airport Seq ID
DestCityMarketID
DestCityName
Dest State
Dest State Fips
DestStateName
Distance Group
DviAirportLandings

## b <- read.csv('2018_Flights.csv')
##  saveRDS('b', file = "201b 8_Flights.Rdata")




Joining 2018 and 2022 data!
First used rbind to join sheets b and c, which contain dfs for 2018 and 2022
Then found 2020 data contained different number of columns, and used -c to drop the two additional columns, which were character vectors on the advertising airline
Converted airline and origin sheets from characters to factors to save space
```{r}
library(dplyr)
b <- readRDS(file = "2022flights.Rdata")
c <- readRDS(file = "2018Flights.Rdata")
help(join)
d <- rbind(b, c)A <- 
View(d)
e <- readRDS(file = "2020flights2.Rdata")
rm(a)
rm(b)
rm(c)
f <- e[-c(33,34)]
g <- rbind(d, f)
rm(e)
library(tibble)
h <- d[-c(21, 23)]
i <- f[-c(21, 23)]
rm(d)
rm(f)
j <- rbind(h, i)
j <- h[-c(21, 23)]
k <- i[-c(21, 23)]
as.factor(j$Airline)
as.factor(k$Airline)
as.factor(j$Origin)
as.factor(k$Origin)
l <- rbind(j, k)
View(j)
j <- j[-c(42)]
k <- k[-c(42)]
saveRDS('l', file = "all_flights.Rdata")












saveRDS(l, file = "all_flights_clean.Rdata")
j <-j[-c(21,22,23,24)]
k <- k[-c(21,22,23,24)]
l <- rbind (j, k)
rm(j)


##figuring out how to get a lone day
library(lubridate)
yday<-(flights$FlightDate)
as.Date(flights$FlightDate)
help(yday)
day(flights$FlightDate)
rm(yday)
as.factor(flights$FlightDate)
flights$flightDate<-as.Date(flights$FlightDate, format = "%m/%d/%Y")
flights <- flights[-c(21,22)]
yday(flights$FlightDate)
rm(yday)
flights <- flights[-c(1)]
as.Date(flights$flightDate) 
day(flights$flightDate)
flights$flight_day <- day(flights$flightDate)
flights <- flights[-c(36)]
flights$year_day <- yday(flights$flightDate)

flights <- readRDS("all_flights_clean.Rdata")
flights2 <- flights[-c(17:37)]
flights2 <- flights2[-c(11:16)]
flights2 <- flights2[-c(5:8)]
flights2$FlightDate<-as.Date(flights2$FlightDate, format = "%m/%d/%Y")
flights2$year_day <- yday(flights2$FlightDate)
flights2$FlightDate<-as.Date(flights2$FlightDate, format = "%m/%d/%Y")
flights2$year_day <- yday(flights2$FlightDate)
saveRDS(flights2, file = "flights_clean_abbreviated.RDS")


##creating three separate clean tables rather than one large one for data restructuring 

a <- readRDS(file = "2018Flights.Rdata")
a <- a[-c(5:8)]
a <- a[-c(7:42)]
a$FlightDate<-as.Date(a$FlightDate, format = "%m/%d/%Y")
a$year_day <- yday(a$FlightDate)
saveRDS(a, file = "2018_clean_concise.RDS")

a <- readRDS(file = "2020flights2.Rdata")
a <- a[-c(5:8)]
a <- a[-c(7:42)]
a <- a[-c(7:8)]
a$FlightDate<-as.Date(a$FlightDate, format = "%m/%d/%Y")
a$year_day <- yday(a$FlightDate)
saveRDS(a, file = "2020_clean_concise.RDS")

a <- readRDS(file = "2022flights.Rdata")
a <- a[-c(5:8)]
a <- a[-c(7:42)]
a$FlightDate<-as.Date(a$FlightDate, format = "%m/%d/%Y")
a$year_day <- yday(a$FlightDate)
saveRDS(a, file = "2022_clean_concise.RDS")

##essential code - creates a vector containing the day of the year of a flight
library(lubridate)
flights$FlightDate<-as.Date(flights$FlightDate, format = "%m/%d/%Y")
flights$year_day <- yday(flights$flightDate)


## airportr package, gives lat and longitude for each airport - AIRPORT MAP DATA WRANGLING
install.packages("airportr")
library(airportr)
library(tidyverse)
airport_location("NAME", input_type = "IATA")
a <- readRDS(file = "all_flights_clean.Rdata")
View(a)
## is the airplane data package vectorized? Try that 
## Merging the data sets for Airports lat and Long
install.packages("airportr")
library(airportr)
View(airports)
install.packages("airportr")
c <- airports
saveRDS(c, file = "airport_information.Rdata")
c <- readRDS(file = "airport_information.Rdata")
View(c)

##data cleanup tidyverse, used filter function to make all airports in the United States and destinations listed in our data set 
library(dplyr)
library(tidyverse)
f <- readRDS(file = "airport_information.Rdata")
a <- readRDS(file = "all_flights_clean.Rdata")
View(a)
View(f)
select(a,Origin)
b <- select(a,Origin)
View(b)
d <- unique(b$Origin)
e <- readRDS(file = "airports_IATA.Rdata")
c %>% filter(IATA %in% e$IATA) 
usairports <- c %>% filter(IATA %in% e$IATA)
View(usairports)

## remove line 9 airports <- readRDS("airport_information.Rdata") from the server and replace with 
airports <- readRDS("usairports.Rdata)

##adding Airport IDs into App 
##server at line 10 
airports <- readRDS("usairportspopups.Rdata")

##at line 64 builds a functional map with popups
library(htmltools)
 output$mymap <- renderLeaflet ({
  leaflet(data = airports[1:17],) %>%
      addTiles() %>%
      addMarkers(lng = ~Longitude, lat= ~Latitude)
  })
  
##adding content to popups @ line 60 builds a functional map with popups
  output$mymap <- renderLeaflet ({
    leaflet(data = airports[1:17],) %>%
      setView(lng =-95.15, lat = 40, zoom = 3.5) %>%
      addTiles() %>%
      addMarkers(lng = ~Longitude, lat= ~Latitude, popup = airports$popups)
  })
  
##merge column function -> paste 
paste("Name:", c$Name, "City:", c$City, "IATA ID:", c$IATA, "Timezone:", c$Timezone) 
c$popups <- paste("Name:", c$Name, "City:", c$City, "IATA ID:", c$IATA, "Timezone:", c$Timezone)
saveRDS(c, file = "usairportspopups.Rdata")
##Final clean popups
c$popupsclean <- paste("Name:", c$Name, "   Iata ID:", c$IATA, "   City:", c$City, "   Timezone:", c$Timezone)
saveRDS(c, file = "usairportspopupsfinal.Rdata")

## cut data into thirds so only one is ever loaded at a time in your dropdown menu (all_flights_clean.Rdata). 
##loaded flights + reactive expression
({
tabletoload <- input$yearselecter
readRDS(tabletoload)
})
output$plot <- renderPlot({
ggplot(loadedflights()) + 
})

##dropdown menu
choices = list(label = value, 
label = value, 
)

##For the data tab, compare flight differences before and after covid-19? 
## wants correrational analysis, certain origins better, certain destination better, what visualizations address those questions


Correlational Graph addition 
Avg U.S. airport delay by time of year 
##scatterplot
group_by(c$FlightDate) %>%
summarise(ave_delay = mean(c$DepDelayMinutes, na.rm = TRUE)) %>%
ggplot(aes(c$FlightDate,c$ave_delay)) + geom_point()

ggplot(c,aes(FlightDate,DepDelayMinutes)) + geom_point() 




rsconnect::writeManifest()
getwd()


